\bf Tearing Modes in Tokamak Plasmas

### 2021-10-25 Documentation/TearingModeTheory.tex
Added to tearing mode theory report

 The dominant terms in the electron and ion momentum conservation equations,
(\ref{e3.151b}) and (\ref{e3.152b}), respectively, yield
\begin{align}\label{e78}
\nabla p_e + n\,e\,({\bf E}+{\bf V}_e\times {\bf B}) &\simeq 0,\\[0.5ex]
\nabla p_i - n\,e\,({\bf E}+{\bf V}_i\times {\bf B}) &\simeq 0.\label{e79}
\end{align}
The sum of the preceding two equations gives
\begin{equation}\label{e80}
 {\bf j}\times {\bf B} \simeq-\nabla p,
\end{equation}
where $p=p_e+p_i$. In other words, a conventional tokamak plasma exists in a state in which electromagnetic
forces exactly balance the  total scalar pressure force. 
It follows from the previous equation that 
\begin{equation}
{\bf B}\cdot\nabla p\simeq 0.
\end{equation}
In other words, the pressure in a conventional tokamak plasma is constant on magnetic flux-surfaces. 
 However, given that $p=n\,(T_e+T_i)$, and making use of Eqs.~(\ref{e75}) and (\ref{e77}), we deduce that 
\begin{equation}
{\bf B}\cdot\nabla n\simeq 0.
\end{equation}
In other words, the particle density in a conventional tokamak plasma is also  constant on
magnetic flux-surfaces.

@@

\section{Cylindrical Tearing Mode Theory}
\subsection{Tokamak Equilibrium}
Consider a large aspect-ratio tokamak plasma equilibrium with whose magnetic flux-surfaces map out (almost) concentric circles in the poloidal plane. 
Let us employ a conventional set of right-handed cylindrical coordinates, $r$, $\theta$, $z$. The
equilibrium magnetic flux-surfaces lie on surfaces of constant $r$. The system is
assumed to be periodic in the $z$ (`toroidal') direction, with periodicity length $2\pi\,R_0$, 
where $R_0$ is the simulated {\em major radius}\/ of the plasma. The {\em safety-factor}\/ profile takes the form
\begin{equation}\label{e90}
q(r)= \frac{r\,B_z}{R_0\,B_\theta(r)},
\end{equation}
where $B_z$ is the constant `toroidal' magnetic field-strength, and $B_\theta(r)$ is
the poloidal magnetic field-strength. It is assumed that $q\sim{\cal O}(1)$. 
The equilibrium `toroidal' current density
satisfies
\begin{equation}
\mu_0\,j_z(r) = \frac{(r\,B_\theta)'}{r},
\end{equation}
where $'$ denotes $d/dr$. Finally, the standard large aspect-ratio
tokamak orderings,  
\begin{equation}\label{e92}
\frac{r}{R_0}\ll 1,~~~ \frac{B_\theta}{B_z}\ll 1,
\end{equation}
are adopted (Fitzpatrick 1993).

\subsection{Perturbed Magnetic Field}
Consider a tearing mode perturbation that has $m$ periods in the poloidal direction,
and $n$ periods in the `toroidal' direction, where $m/n\sim {\cal O}(1)$.  We can write the perturbed magnetic field in the form
\begin{equation}\label{e93}
\delta{\bf B} \simeq \nabla\times (\psi\,{\bf e}_z),
\end{equation}
where
\begin{equation}\label{e93a}
\psi(r,\theta,\phi,t)= \psi(r,t)\,\exp[{\rm i}\,(m\,\theta-n\,\phi)].
\end{equation}
Here, $\phi=z/R_0$ is a simulated toroidal angle. The perturbed current density becomes
\begin{equation}\label{e94}
\mu_0\,\delta {\bf j} \simeq - \nabla^2\psi\,{\bf e}_z.
\end{equation}

\subsection{Cylindrical Tearing Mode Equation}
The tearing perturbation is governed by the linearized version of Equation~(\ref{e84}), which yields
\begin{equation}\label{e96}
(\delta{\bf B}\cdot \nabla)\,{\bf j} + ({\bf B}\cdot\nabla)\,\delta{\bf j} - (\delta{\bf j}\cdot\nabla)\,{\bf B} - 
({\bf j}\cdot\nabla)\,\delta {\bf B}\simeq {\bf 0}.
\end{equation}
Making use of Equations~(\ref{e90})--(\ref{e94}), the $z$-component of the previous equation gives
the {\em cylindrical tearing mode equation}\/ (Fitzpatrick 1993),
\begin{equation}
\frac{\partial^2\psi}{dr^2} + \frac{1}{r}\,\frac{\partial\psi}{\partial r}-\frac{m^2}{r^2}\,\psi - \frac{J_z'\,\psi}{r\,(1/q-n/m)}\simeq 0,
\end{equation}
where 
\begin{equation}
J_z(r)= \frac{R_0\,\mu_0\,j_z(r)}{B_z}.
\end{equation}

@@

Note that Equation~(\ref{e100}) is singular at the so-called {\em resonant}\/ magnetic flux-surface, radius $r=r_s$, at which 
\begin{equation}\label{e104}
q(r_s)= \frac{m}{n}.
\end{equation}
At the resonant surface, ${\bf k}\cdot{\bf B}=0$, where ${\bf k} = (0,\,m/r,\,-n/R_0)$ is the wavevector of the tearing perturbation. 
The value of $\psi(r,t)$ at the resonant surface---namely, ${\mit\Psi}_s(t)$---is known as the {\em reconnected magnetic flux}. 
Note that ${\mit\Psi}_s$ is, in general, a complex quantity. 
Let $\rho=(r-r_s)/r_s$. 
The solution of Equation~(\ref{e100}) in the vicinity of the resonant surface is
\begin{equation}\label{e105}
\hat{\psi}_s(\rho) = 1 + {\mit\Delta}_{s+}\,\rho + \alpha_s\,\rho\,\ln|\rho| + {\cal O}(\rho^2)
\end{equation}
for $\rho>0$, and
\begin{equation}\label{e106}
\hat{\psi}_s(\rho) = 1 + {\mit\Delta}_{s-}\,\rho + \alpha_s\,\rho\,\ln|\rho| + {\cal O}(\rho^2)
\end{equation}
for $\rho<0$. Here, 
\begin{align}
\alpha_s& =- \left(\frac{r\,q\,J_z'}{s}\right)_{r=r_s},\\[0.5ex]
s(r) &= \frac{r\,q'}{q}.
\end{align}
Moreover,  the real parameters ${\mit\Delta}_{s+}$ and  ${\mit\Delta}_{s-}$ are fully determined by Equation~(\ref{e100})
and the boundary conditions (\ref{e101})--(\ref{e103}). Note that, in general, $\psi$ is continuous across the resonant surface,
whereas $\partial\psi/\partial r$ is discontinuous. The discontinuity in $\partial \psi/\partial r$ implies the presence of a radially
thin current sheet  at the resonant surface. This current sheet can only be resolved by retaining more terms in the perturbed plasma equation of motion, (\ref{e96}).

@@

In general, $\psi$ is continuous across the wall, whereas $\partial\psi/\partial r$ is discontinuous. The discontinuity in $\partial\psi/\partial r$ is caused by the currents induced in the wall. The complex quantity ${\mit\Psi}_w(t)$ determines the amplitude and
phase of the perturbed magnetic flux that penetrates the wall. The complex quantity
\begin{equation}\label{e118}
{\mit\Delta\Psi}_w = \left[r\,\frac{\partial\psi}{\partial r}\right]_{r_{w-}}^{r_{w+}}
\end{equation}
parameterizes the amplitude and phase of the current sheet flowing in the wall.
Simultaneously matching the outer solution (\ref{e111}) across the resonant surface and the wall yields
\begin{align}
{\mit\Delta\Psi}_s &= E_{ss}\,{\mit\Psi}_s + E_{sw}\,{\mit\Psi}_w,\\[0.5ex]
{\mit\Delta\Psi}_w &= E_{ws}\,{\mit\Psi}_s+ E_{ww}\,{\mit\Psi}_w
\end{align}
(Fitzpatrick 1993).
Here,
\begin{align}
E_{ww}&= \left[r\,\frac{d\hat{\psi}_w}{dr}\right]_{r_{w-}}^{r_{w+}},\\[0.5ex]
E_{sw} &=\left[r\,\frac{{d\hat\psi}_w}{dr}\right]_{r=r_{s+}},\label{e121}\\[0.5ex]
E_{ws} &=-\left[r\,\frac{{d\hat\psi}_s}{dr}\right]_{r=r_{w-}}\label{e122}
\end{align}
are real quantities determined by the solution in the outer region.

@@

\subsection{Resistive Wall Physics}
It is clear from Equation~(\ref{e93}) that
\begin{equation}
\delta {\bf A} \simeq \psi\,{\bf e}_z,
\end{equation}
where $\delta {\bf A}$ is the perturbed magnetic vector potential. 
Hence, the perturbed electric field can be written
\begin{equation}
\delta {\bf E} \simeq -\nabla\delta{\mit\Phi} - \frac{\partial\psi}{\partial t}\,{\bf e}_z,
\end{equation}
where $\delta{\mit\Phi}$ is the perturbed electric scalar potential. 
Assuming that $|\nabla\delta{\mit\Phi}|\sim |\partial\psi/\partial t|$, the $z$-component of the previous
equation yields
\begin{equation}
\delta E_z\simeq -\frac{\partial\psi}{\partial t},
\end{equation}
where use has been made of the large aspect-ratio orderings (\ref{e92}). Within the wall, we
can write
\begin{equation}
\delta E_z \simeq -\frac{d{\mit\Psi}_w}{dt}.
\end{equation}
Here, we are making use of the so-called {\em thin wall approximation}, according to which $\psi$ is assumed to only vary weakly in $r$ across
the wall. Ohm's law implies that
\begin{equation}
\delta j_z = \frac{\delta E_z}{\eta_w} = -\frac{1}{\eta_w}\,\frac{d{\mit\Psi}_w}{dt}
\end{equation}
within the wall, where $\eta_w$ is the wall resistivity. Equations~(\ref{e94}) and (\ref{e118}) yield
\begin{equation}\label{e132}
{\mit\Delta\Psi}_w = -\mu_0\int_{r_{w-}}^{r_{w+}}r\,\delta j_z\,dr.
\end{equation}
Suppose that the radial thickness of the wall is $\delta_w\ll r_w$. The previous two equations yield
\begin{equation}\label{e133}
{\mit\Delta\Psi}_w = \tau_w\,\frac{d{\mit\Psi}_w}{dt},
\end{equation}
where
\begin{equation}
\tau_w = \frac{\mu_0\,r_w\,\delta_w}{\eta_w}
\end{equation}
is the so-called {\em time constant}\/ of the wall (Fitzpatrick 1993). The thin wall approximation is
valid as long as
\begin{equation}
\frac{r_w}{\delta_w}\gg \tau_w\left|\frac{d\ln{\mit\Psi}_w}{dt}\right|.
\end{equation}

\subsection{Resonant Layer Physics}
By analogy with Equation~(\ref{e133}), we can write
\begin{equation}
{\mit\Delta\Psi}_s = \tau_s\left(\frac{d{\mit\Psi}_w}{dt} + {\rm i}\,\varpi_s\,{\mit\Psi}_s\right).
\end{equation}
Here, $\tau_s$ is the {\em reconnection time}, whereas $\varpi_s$ is the so-called {\em natural frequency}. 
The natural frequency is the phase velocity of the tearing mode when the wall is perfectly conducting. The
natural frequency is non-zero because the reconnected flux in the resonant layer is convected by the local plasma
flow (Fitzpatrick 1993).

\subsection{Solution in Presence of External Magnetic Field-Coil}
Suppose that the perfectly-conducting wall at $r=r_c$ is replaced by a magnetic field-coil that carries a
helical current possesing $m$ periods in the poloidal direction, and $n$ periods in the `toroidal' direction. Let the current density in the coil
take the form
\begin{equation}\label{e135}
\delta j_z = -\frac{I_c(t)}{r_c}\,\delta(r-r_c)\,{\rm e}^{\,{\rm i}\,(m\,\theta-n\,\phi)}.
\end{equation}
Here, the complex quantity $I_c(t)$ parameterizes the amplitude and phase of the helical current flowing in the field-coil.

@@

In general, $\psi$ is continuous across the field-coil, whereas $\partial\psi/\partial r$ is discontinuous. The discontinuity in $\partial\psi/\partial r$ is caused by the currents flowing in the coil. The complex quantity ${\mit\Psi}_c(t)$ determines the amplitude and
phase of the perturbed magnetic flux at the coil. The complex quantity
\begin{equation}\label{e118a}
{\mit\Delta\Psi}_c = \left[r\,\frac{\partial\psi}{\partial r}\right]_{r_{c-}}^{r_{c+}}
\end{equation}
parameterizes the amplitude and phase of the current sheet flowing in the coil.
It follows from Equations~(\ref{e93a}), (\ref{e94}), and (\ref{e135})  that
\begin{equation}
{\mit\Delta\Psi}_c = \mu_0\,I_c.
\end{equation}

@@

\subsection{Electromagnetic Torques}
The flux-surface integrated poloidal and `toroidal' electromagnetic torque densities acting on the plasma
can be written
\begin{align}
T_\theta(r) &= \oint\oint R_0\,r^{\,2}\,{\bf j}\times {\bf B}\cdot {\bf e}_\theta\,d\theta\,d\phi,\\[0.5ex]
T_\phi(r) &= \oint\oint R_0^{\,2}\,r\,{\bf j}\times {\bf B}\cdot {\bf e}_z\,d\theta\,d\phi,
\end{align}
respectively. However, according to Equation~(\ref{e80}),
\begin{equation}\label{e150}
{\bf j}\times {\bf B} \simeq -\nabla p.
\end{equation}
Given that the scalar pressure is a single-valued function of $\theta$ and $\phi$, it immediately follows
that $T_\theta=T_\phi = 0$ throughout the plasma. The only exception to this rule occurs in the immediate vicinity of the
resonant surface, where Equation~(\ref{e150}) breaks down. It follows that we can write
\begin{align}
T_\theta(r) &= T_{\theta\,s}\,\delta(r-r_s),\\[0.5ex]
T_\phi(r) &= T_{\phi\,s}\,\delta(r-r_s),
\end{align}
where
\begin{align}\label{e153}
T_{\theta\,s} &=  \frac{1}{4}\int_{r_{s-}}^{r_{s+}}\oint\oint R_0\,r^{\,2}\,(\delta j_z\,\delta B_r^{\,\ast} + \delta j_z^{\,\ast}\,\delta B_r)\,dr\,d\theta\,d\phi,\\[0.5ex]
T_{\phi\,s} &=  -\frac{1}{4}\int_{r_{s-}}^{r_{s+}}\oint\oint R_0^{\,2}\,r\,(\delta j_\theta\,\delta B_r^{\,\ast} + \delta j_\theta^{\,\ast}\,\delta B_r)\,dr\,d\theta\,d\phi
\end{align}
are the net poloidal and `toroidal' torques, respectively, acting at the resonant surface. 
Note that the equilibrium torques are zero because $B_r=0$. Furthermore, the linear (in perturbed quantities) torques 
average to zero over the flux-surface. Hence, the largest non-zero torques  are quadratic in perturbed quantities. 
According to Equation~(\ref{e82}), 
\begin{equation}
\nabla \cdot \delta {\bf j} = {\rm i}\left(\frac{m}{r}\,\delta j_\theta - \frac{n}{R_0}\,\delta j_z\right) = 0,
\end{equation}
where $\delta{\bf j}$ is the density of the current sheet flowing at the resonant surface, and we have made use
of the fact that $\delta j_r\simeq 0$. 
 It follows from the previous three equations that
\begin{equation}\label{e156}
T_{\phi\,s} = - \frac{n}{m}\,T_{\theta\,s}.
\end{equation}
Finally, Equations~(\ref{e93}), (\ref{e94}), (\ref{e153}), and (\ref{e156}) imply that
\begin{align}\label{e157}
T_{\theta\,s} &= -\frac{2\pi^2\,R_0\,m}{\mu_0}\,{\rm Im}({\mit\Delta\Psi}_s\,{\mit\Psi}_s^{\,\ast}),\\[0.5ex]
T_{\phi\,s} &= \frac{2\pi^2\,R_0\,n}{\mu_0}\,{\rm Im}({\mit\Delta\Psi}_s\,{\mit\Psi}_s^{\,\ast})
\end{align}
(Fitzpatrick 1993).

>>>>

 The dominant terms in the electron and ion momentum conservation equations,
(\ref{e3.151b}) and (\ref{e3.152b}), respectively, yield
\begin{align}\label{e78}
\nabla p_e + n\,e\,({\bf E}+{\bf V}_e\times {\bf B}) &\simeq 0,\\[0.5ex]
\nabla p_i - n\,e\,({\bf E}+{\bf V}_i\times {\bf B}) &\simeq 0.\label{e79}
\end{align}
The sum of the preceding two equations gives
\begin{equation}\label{e80}
 {\bf j}\times {\bf B} \simeq\nabla p,
\end{equation}
where $p=p_e+p_i$. In other words, a conventional tokamak plasma exists in a state in which electromagnetic
forces exactly balance the  total scalar pressure force. 
It follows from the previous equation that 
\begin{equation}
{\bf B}\cdot\nabla p\simeq 0.
\end{equation}
In other words, the pressure in a conventional tokamak plasma is constant on magnetic flux-surfaces. 
 However, given that $p=n\,(T_e+T_i)$, and making use of Eqs.~(\ref{e75}) and (\ref{e77}), we deduce that 
\begin{equation}
{\bf B}\cdot\nabla n\simeq 0.
\end{equation}
In other words, the particle density in a conventional tokamak plasma is also  constant on
magnetic flux-surfaces.

@@

\section{Cylindrical Tearing Mode Theory}
\subsection{Tokamak Equilibrium}
Consider a large aspect-ratio, toroidal,  tokamak plasma equilibrium whose magnetic flux-surfaces map out (almost) concentric circles in the poloidal plane. 
Let us employ a conventional set of right-handed cylindrical coordinates, $r$, $\theta$, $z$. The
equilibrium magnetic flux-surfaces lie on surfaces of constant $r$. The system is
assumed to be periodic in the $z$ (`toroidal') direction, with periodicity length $2\pi\,R_0$, 
where $R_0$ is the simulated {\em major radius}\/ of the plasma. The {\em safety-factor}\/ profile takes the form
\begin{equation}\label{e90}
q(r)= \frac{r\,B_z}{R_0\,B_\theta(r)},
\end{equation}
where $B_z$ is the constant `toroidal' magnetic field-strength, and $B_\theta(r)$ is
the poloidal magnetic field-strength. It is assumed that $q\sim{\cal O}(1)$. 
The equilibrium `toroidal' current density
satisfies
\begin{equation}
\mu_0\,j_z(r) = \frac{(r\,B_\theta)'}{r},
\end{equation}
where $'$ denotes $d/dr$. Finally, the standard large aspect-ratio
tokamak orderings,  
\begin{equation}\label{e92}
\frac{r}{R_0}\ll 1,~~~ \frac{B_\theta}{B_z}\ll 1,
\end{equation}
are adopted (Fitzpatrick 1993).

\subsection{Perturbed Magnetic Field}
Consider a tearing mode perturbation that has $m$ periods in the poloidal direction,
and $n$ periods in the `toroidal' direction, where $m/n\sim {\cal O}(1)$.  We can write the perturbed magnetic field in the form
\begin{equation}\label{e93}
\delta{\bf B} \simeq \nabla\times (\psi\,{\bf e}_z),
\end{equation}
where
\begin{equation}\label{e93a}
\psi(r,\theta,\phi,t)= \psi(r,t)\,\exp[{\rm i}\,(m\,\theta-n\,\phi)].
\end{equation}
Here, $\phi=z/R_0$ is a simulated toroidal angle. The perturbed current density becomes
\begin{equation}\label{e94}
\mu_0\,\delta {\bf j} \simeq - \nabla^2\psi\,{\bf e}_z.
\end{equation}
It is assumed that $|\delta{\bf B}|\ll |{\bf B}|$.

\subsection{Cylindrical Tearing Mode Equation}
To lowest order, the tearing perturbation is governed by the linearized version of Equation~(\ref{e84}), which yields
\begin{equation}\label{e96}
(\delta{\bf B}\cdot \nabla)\,{\bf j} + ({\bf B}\cdot\nabla)\,\delta{\bf j} - (\delta{\bf j}\cdot\nabla)\,{\bf B} - 
({\bf j}\cdot\nabla)\,\delta {\bf B}\simeq {\bf 0}.
\end{equation}
Making use of Equations~(\ref{e90})--(\ref{e94}), the $z$-component of the previous equation gives
the {\em cylindrical tearing mode equation}\/ (Fitzpatrick 1993),
\begin{equation}
\frac{\partial^2\psi}{dr^2} + \frac{1}{r}\,\frac{\partial\psi}{\partial r}-\frac{m^2}{r^2}\,\psi - \frac{J_z'\,\psi}{r\,(1/q-n/m)}\simeq 0,
\end{equation}
where 
\begin{equation}
J_z(r)= \frac{R_0\,\mu_0\,j_z(r)}{B_z}.
\end{equation}

@@

Note that Equation~(\ref{e100}) is singular at the so-called {\em resonant}\/ magnetic flux-surface, radius $r=r_s$, at which 
\begin{equation}\label{e104}
q(r_s)= \frac{m}{n}.
\end{equation}
At the resonant surface, ${\bf k}\cdot{\bf B}=0$, where ${\bf k} = (0,\,m/r,\,-n/R_0)$ is the wavevector of the tearing perturbation. 
The value of $\psi(r,t)$ at the resonant surface---namely, ${\mit\Psi}_s(t)$---is known as the {\em reconnected magnetic flux}. 
Note that ${\mit\Psi}_s$ is, in general, a complex quantity. 
Let $\rho=(r-r_s)/r_s$. 
The solution of Equation~(\ref{e100}) in the vicinity of the resonant surface is
\begin{equation}\label{e105}
\hat{\psi}_s(\rho) = 1 + {\mit\Delta}_{s+}\,\rho + \alpha_s\,\rho\,\ln|\rho| + {\cal O}(\rho^2)
\end{equation}
for $\rho>0$, and
\begin{equation}\label{e106}
\hat{\psi}_s(\rho) = 1 + {\mit\Delta}_{s-}\,\rho + \alpha_s\,\rho\,\ln|\rho| + {\cal O}(\rho^2)
\end{equation}
for $\rho<0$. Here, 
\begin{align}
\alpha_s& =- \left(\frac{r\,q\,J_z'}{s}\right)_{r=r_s},\\[0.5ex]
s(r) &= \frac{r\,q'}{q}.
\end{align}
Moreover,  the real parameters ${\mit\Delta}_{s+}$ and  ${\mit\Delta}_{s-}$ are fully determined by Equation~(\ref{e100})
and the boundary conditions (\ref{e101})--(\ref{e103}). Note that, in general, $\psi$ is continuous across the resonant surface (in accordance with Maxwell's equations),
whereas $\partial\psi/\partial r$ is discontinuous. The discontinuity in $\partial \psi/\partial r$ implies the presence of a radially
thin, helical,  current sheet  at the resonant surface. This current sheet can only be resolved by retaining more terms in the perturbed plasma equation of motion, (\ref{e96}).

@@

In general, $\psi$ is continuous across the wall (in accordance with Maxwell's equations), whereas $\partial\psi/\partial r$ is discontinuous. The discontinuity in $\partial\psi/\partial r$ is caused by the currents induced in the wall. The complex quantity ${\mit\Psi}_w(t)$ determines the amplitude and
phase of the perturbed magnetic flux that penetrates the wall. The complex quantity
\begin{equation}\label{e118}
{\mit\Delta\Psi}_w = \left[r\,\frac{\partial\psi}{\partial r}\right]_{r_{w-}}^{r_{w+}}
\end{equation}
parameterizes the amplitude and phase of the helical current sheet flowing in the wall.
Simultaneously matching the outer solution (\ref{e111}) across the resonant surface and the wall yields
\begin{align}
{\mit\Delta\Psi}_s &= E_{ss}\,{\mit\Psi}_s + E_{sw}\,{\mit\Psi}_w,\\[0.5ex]
{\mit\Delta\Psi}_w &= E_{ws}\,{\mit\Psi}_s+ E_{ww}\,{\mit\Psi}_w
\end{align}
(Fitzpatrick 1993).
Here,
\begin{align}
E_{ww}&= \left[r\,\frac{d\hat{\psi}_w}{dr}\right]_{r_{w-}}^{r_{w+}},\\[0.5ex]
E_{sw} &=\left[r\,\frac{{d\hat\psi}_w}{dr}\right]_{r=r_{s+}},\label{e121}\\[0.5ex]
E_{ws} &=-\left[r\,\frac{{d\hat\psi}_s}{dr}\right]_{r=r_{w-}}\label{e122}
\end{align}
are real quantities determined by the solution in the outer region.

@@

\subsection{Resistive Wall Physics}
It is clear from Equation~(\ref{e93}) that
\begin{equation}
\delta {\bf A} \simeq \psi\,{\bf e}_z,
\end{equation}
where $\delta {\bf A}$ is the perturbed magnetic vector potential. 
Hence, the perturbed electric field can be written
\begin{equation}
\delta {\bf E} \simeq -\nabla\delta{\mit\Phi} - \frac{\partial\psi}{\partial t}\,{\bf e}_z,
\end{equation}
where $\delta{\mit\Phi}$ is the perturbed electric scalar potential. 
Assuming that $|\nabla\delta{\mit\Phi}|\sim |\partial\psi/\partial t|$, the $z$-component of the previous
equation yields
\begin{equation}
\delta E_z\simeq -\frac{\partial\psi}{\partial t},
\end{equation}
where use has been made of the large aspect-ratio orderings (\ref{e92}). Within the wall, we
can write
\begin{equation}
\delta E_z \simeq -\frac{d{\mit\Psi}_w}{dt}.
\end{equation}
Here, we are making use of the so-called {\em thin wall approximation}, according to which $\psi$ is assumed to only vary weakly in $r$ across
the wall. Ohm's law implies that
\begin{equation}
\delta j_z = \frac{\delta E_z}{\eta_w} = -\frac{1}{\eta_w}\,\frac{d{\mit\Psi}_w}{dt}
\end{equation}
within the wall, where $\eta_w$ is the wall resistivity. Equations~(\ref{e94}) and (\ref{e118}) yield
\begin{equation}\label{e132}
{\mit\Delta\Psi}_w = -\mu_0\int_{r_{w-}}^{r_{w+}}r\,\delta j_z\,dr.
\end{equation}
Suppose that the radial thickness of the wall is $\delta_w\ll r_w$. The previous two equations give
\begin{equation}\label{e133}
{\mit\Delta\Psi}_w = \tau_w\,\frac{d{\mit\Psi}_w}{dt},
\end{equation}
where
\begin{equation}
\tau_w = \frac{\mu_0\,r_w\,\delta_w}{\eta_w}
\end{equation}
is the so-called {\em time constant}\/ of the wall (Fitzpatrick 1993). The thin wall approximation is
valid as long as
\begin{equation}
\frac{r_w}{\delta_w}\gg \tau_w\left|\frac{d\ln{\mit\Psi}_w}{dt}\right|.
\end{equation}

\subsection{Resonant Layer Physics}
By analogy with Equation~(\ref{e133}), we can write
\begin{equation}
{\mit\Delta\Psi}_s = \tau_s\left(\frac{d{\mit\Psi}_w}{dt} + {\rm i}\,\varpi_s\,{\mit\Psi}_s\right).
\end{equation}
Here, $\tau_s$ is the {\em reconnection time}\/ (i.e., the typical timescale on which magnetic reconnection takes place in the
resonant layer surrounding the resonant surface), whereas $\varpi_s$ is the so-called {\em natural frequency}. 
The natural frequency is the phase velocity of the tearing mode when the wall is perfectly conducting. The
natural frequency is non-zero because the reconnected flux in the resonant layer is convected by the local plasma
flow (Fitzpatrick 1993).

\subsection{Solution in Presence of External Magnetic Field-Coil}
Suppose that the perfectly-conducting wall at $r=r_c$ is replaced by a magnetic field-coil that carries a
helical current possesing $m$ periods in the poloidal direction, and $n$ periods in the `toroidal' direction. Let the current density in the field-coil
take the form
\begin{equation}\label{e135}
\delta j_z = -\frac{I_c(t)}{r_c}\,\delta(r-r_c)\,{\rm e}^{\,{\rm i}\,(m\,\theta-n\,\phi)}.
\end{equation}
Here, the complex quantity $I_c(t)$ parameterizes the amplitude and phase of the helical current flowing in the field-coil.

@@

In general, $\psi$ is continuous across the field-coil (in accordance with Maxwell's equations), whereas $\partial\psi/\partial r$ is discontinuous. The discontinuity in $\partial\psi/\partial r$ is caused by the currents flowing in the coil. The complex quantity ${\mit\Psi}_c(t)$ determines the amplitude and
phase of the perturbed magnetic flux at the coil. The complex quantity
\begin{equation}\label{e118a}
{\mit\Delta\Psi}_c = \left[r\,\frac{\partial\psi}{\partial r}\right]_{r_{c-}}^{r_{c+}}
\end{equation}
parameterizes the amplitude and phase of the current sheet flowing in the coil.
It follows from Equations~(\ref{e93a}), (\ref{e94}), and (\ref{e135})  that
\begin{equation}
{\mit\Delta\Psi}_c = \mu_0\,I_c.
\end{equation}

@@

\subsection{Electromagnetic Torques}\label{storque}
The flux-surface integrated poloidal and `toroidal' electromagnetic torque densities acting on the plasma
can be written
\begin{align}\label{e152}
T_\theta(r) &= \left\langle r\,{\bf j}\times {\bf B}\cdot {\bf e}_\theta\right\rangle\\[0.5ex]
T_\phi(r) &= \left\langle R_0\,{\bf j}\times {\bf B}\cdot {\bf e}_z\right\rangle,
\end{align}
respectively, 
where
\begin{equation}
\left\langle\cdots\right\rangle \equiv \oint\oint r\,R_0\,(\cdots)\,d\theta\,d\phi
\end{equation}
is the flux-surface average operator. 
 However, according to Equation~(\ref{e80}),
\begin{equation}\label{e150}
{\bf j}\times {\bf B} \simeq \nabla p.
\end{equation}
Given that the scalar pressure is a single-valued function of $\theta$ and $\phi$, it immediately follows
that $T_\theta=T_\phi = 0$ throughout the plasma. The only exception to this rule occurs in the immediate vicinity of the
resonant surface, where Equation~(\ref{e150}) breaks down. It follows that we can write
\begin{align}\label{e156a}
T_\theta(r) &= T_{\theta\,s}\,\delta(r-r_s),\\[0.5ex]
T_\phi(r) &= T_{\phi\,s}\,\delta(r-r_s),
\end{align}
where
\begin{align}\label{e153}
T_{\theta\,s} &=  \frac{1}{4}\int_{r_{s-}}^{r_{s+}}\oint\oint R_0\,r^{\,2}\,(\delta j_z\,\delta B_r^{\,\ast} + \delta j_z^{\,\ast}\,\delta B_r)\,dr\,d\theta\,d\phi,\\[0.5ex]
T_{\phi\,s} &=  -\frac{1}{4}\int_{r_{s-}}^{r_{s+}}\oint\oint R_0^{\,2}\,r\,(\delta j_\theta\,\delta B_r^{\,\ast} + \delta j_\theta^{\,\ast}\,\delta B_r)\,dr\,d\theta\,d\phi
\end{align}
are the net poloidal and `toroidal' torques, respectively, acting at the resonant surface. 
Note that the equilibrium torques are zero because $B_r=0$. Furthermore, the linear (in perturbed quantities) torques 
average to zero over the flux-surface. Hence, the largest non-zero torques  are quadratic in perturbed quantities. 
According to Equation~(\ref{e82}), 
\begin{equation}
\nabla \cdot \delta {\bf j} = {\rm i}\left(\frac{m}{r}\,\delta j_\theta - \frac{n}{R_0}\,\delta j_z\right) = 0,
\end{equation}
where $\delta{\bf j}$ is the density of the current sheet flowing at the resonant surface, and we have made use
of the fact that $\delta j_r\simeq 0$. 
 It follows from the previous three equations that
\begin{equation}\label{e156}
T_{\phi\,s} = - \frac{n}{m}\,T_{\theta\,s}.
\end{equation}
Finally, Equations~(\ref{e93}), (\ref{e94}), (\ref{e153}), and (\ref{e156}) imply that
\begin{align}\label{e157}
T_{\theta\,s} &= -\frac{2\pi^2\,R_0\,m}{\mu_0}\,{\rm Im}({\mit\Delta\Psi}_s\,{\mit\Psi}_s^{\,\ast}),\\[0.5ex]
T_{\phi\,s} &= \frac{2\pi^2\,R_0\,n}{\mu_0}\,{\rm Im}({\mit\Delta\Psi}_s\,{\mit\Psi}_s^{\,\ast})
\end{align}
(Fitzpatrick 1993).
